---
title: "template"
output: html_document
date: '2023-10-07'
---

```{r setup, include=FALSE}

```

## R Markdown


## read dataset do cleaning using formula and neglect IDNO DENSITY and ADIPOSITY (reason it can be directly calculated from height and weight) 

```{r read data}
# Load necessary libraries

library(tidyverse)
library(lmtest)
library(ggplot2)

# Assuming df is your data frame
BodyFat <- read.csv("C:/Users/24788/Downloads/BodyFat.csv")
df <- BodyFat

# Calculate the percentage of body fat using the provided formula
df$calculated_bodyfat <- 495 / df$DENSITY - 450

# let's consider values below 2 or above 100 as impossible.
impossible_values <- which(df$calculated_bodyfat <2 | df$calculated_bodyfat > 50|df$BODYFAT<2|df$BODYFAT>50)

# You can then decide to remove these impossible records.
df_clean <- df[-impossible_values, ]


df<-df_clean
#removed unwanted columns like IDNO, DENSITY, ADIPOSITY
df <- df %>%
  select(-IDNO, -DENSITY, -ADIPOSITY, -calculated_bodyfat)

```


## using corrleation matrix to pick up the most related variables,from the plot we see many features are related so we decided to use vif and recursive elimation to further modify out model.

```{r fig.width=14, fig.height=8}

library(ggcorrplot)
library(dplyr)
cor_matrix <- cor(df)
ggcorrplot(cor_matrix, hc.order = TRUE, type = "lower", lab = TRUE)

cor_matrix <- cor(df )

# Extract correlations with BODYFAT
correlations_with_BODYFAT <- cor_matrix[, "BODYFAT"]

# Remove BODYFAT's correlation with itself (which will be 1)
correlations_with_BODYFAT <- correlations_with_BODYFAT[!names(correlations_with_BODYFAT) %in% "BODYFAT"]

# Sort by the absolute value to see the strongest relationships
sorted_correlations <- correlations_with_BODYFAT[order(-abs(correlations_with_BODYFAT))]

# Visualize the sorted correlations using ggplot2
ggplot(data.frame(Feature = names(sorted_correlations), Correlation = sorted_correlations), aes(x=reorder(Feature, Correlation), y=abs(Correlation), fill=Correlation > 0)) +
  geom_bar(stat="identity") +
  coord_flip() +
  scale_fill_manual(values = c("#FF9999", "steelblue"), name = "Correlation", breaks = c(TRUE, FALSE), labels = c("Positive", "Negative")) +
  ggtitle("Correlations with BODYFAT Without DENSITY") +
  xlab("Features") +
  ylab("Pearson Correlation Coefficient ") +
  theme_minimal()
```

## final result and comparion base on rmse and r squared

```{r combined rfe and vif}
library(caret)
library(ggplot2)
library(dplyr)

# Calculate correlations of each variable with BODYFAT
correlations <- cor(df)
target_correlations <- correlations[, "BODYFAT"]

# Select variables that meet the threshold for correlation
highly_correlated_vars <- names(target_correlations[which(abs(target_correlations) > 0.5)])

# Remove BODYFAT from the predictor dataset since it's the target variable
predictor_vars <- highly_correlated_vars[highly_correlated_vars != "BODYFAT"]

# Prepare data for rfe
predictors <- df[, predictor_vars]
response <- df$BODYFAT

set.seed(123) # for reproducibility

# Define training control
train_control <- rfeControl(
  functions = lmFuncs,
  method = "cv",
  number = 10 # number of folds in cross-validation
)

# Run the recursive feature elimination
results <- rfe(
  x = predictors, 
  y = response,
  sizes = c(1:length(predictor_vars)), # to test models with all possible numbers of predictors
  rfeControl = train_control
)

# Extract the results
result_frame <- data.frame(
  Predictors = results$results$Variables, 
  Rsquared = results$results$Rsquared
)

# Create the plot
ggplot(result_frame, aes(x = Predictors, y = Rsquared)) + 
  geom_point() +
  geom_line() +
  labs(title = "Number of Predictors vs R-squared",
       x = "Number of Predictors",
       y = "R-squared") +
  theme_minimal()


# Print the summary of the RFE result
print(results)

model_simplified <- lm(BODYFAT ~ ABDOMEN + THIGH + HIP + WEIGHT, data = df)

# Check the summary
summary(model_simplified)

# Calculate VIFs for the new model
library(car)
vif_simplified <- vif(model_simplified)
print(vif_simplified)


# Removing the variable with the highest VIF

model_simplified <- lm(BODYFAT ~ ABDOMEN + THIGH + WEIGHT, data = df)

# Check the summary
summary(model_simplified)

# Calculate VIFs for the new model
vif_simplified <- vif(model_simplified)
print(vif_simplified)

# Removing the variable with the highest VIF

model_simplified <- lm(BODYFAT ~ ABDOMEN + THIGH, data = df)

# Check the summary
summary(model_simplified)

# Calculate VIFs for the new model
vif_simplified <- vif(model_simplified)
print(vif_simplified)

plot(model_simplified)
```

```{r test rmse}
# Load necessary library
library(caret)

# Prepare data
set.seed(123)  # for reproducibility
training_index <- createDataPartition(df$BODYFAT, p = 0.8, list = FALSE)
training_data <- df[training_index, ]
testing_data <- df[-training_index, ]

# Fit the full model
full_model <- lm(BODYFAT ~ ., data = training_data)  # replace '.' with actual predictors if you haven't removed 'IDNO' and 'DENSITY'

# Fit the reduced model, based on the variables you've chosen through Pearson's correlation, recursive feature elimination, and VIF
reduced_model <- lm(BODYFAT ~ ABDOMEN + THIGH + WEIGHT, data = training_data)  # replace with your actual formula if different

# Predict on the test set
predictions_full <- predict(full_model, newdata = testing_data)
predictions_reduced <- predict(reduced_model, newdata = testing_data)

# Calculate the RMSE
rmse_full <- sqrt(mean((predictions_full - testing_data$BODYFAT)^2))
rmse_reduced <- sqrt(mean((predictions_reduced - testing_data$BODYFAT)^2))

# Print the results
cat("Full model RMSE:", rmse_full, "\n")
cat("Reduced model RMSE:", rmse_reduced, "\n")

# Define the models
formula1 <- BODYFAT ~ ABDOMEN + THIGH + WEIGHT + HIP
formula2 <- BODYFAT ~ ABDOMEN + THIGH

# Fit the models on the training data
model1 <- lm(formula1, data = training_data)
model2 <- lm(formula2, data = training_data)

# Predict on the testing data
predictions1 <- predict(model1, newdata = testing_data)
predictions2 <- predict(model2, newdata = testing_data)

# Calculate RMSE
rmse1 <- sqrt(mean((predictions1 - testing_data$BODYFAT)^2))
rmse2 <- sqrt(mean((predictions2 - testing_data$BODYFAT)^2))

# Print the RMSE
print(paste("RMSE for model 1 (ABDOMEN + THIGH + WEIGHT + HIP):", rmse1))
print(paste("RMSE for model 2 (ABDOMEN + THIGH):", rmse2))
print(paste("RMSE for model 1 (ABDOMEN + THIGH + WEIGHT ):", rmse_reduced))

print(paste("RMSE for model 1 (full model ):", rmse_full))




```
